---
title: "Getting started: fitting a model to a bundled dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started: fitting a model to a bundled dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(cogzoo)
```

Every model in the zoo shares the same four generics: `cm_simulate()`,
`cm_loglik()`, `cm_fit()`, and `cm_predict()`. This walks through all four
using the Generalized Context Model (`gcm()`, Nosofsky 1986) and its bundled
example dataset.

## What's available

```{r}
list_models()[["gcm"]][c("domain", "task_types", "description")]
```

## Loading a bundled dataset

Datasets ship under `inst/extdata/` with a validated metadata sidecar —
`read_cogdata()` checks the sidecar before returning the data, so a
malformed dataset fails loudly instead of silently.

```{r}
responses <- read_cogdata(system.file(
  "extdata", "nosofsky_1989", "nosofsky_1989_responses.csv",
  package = "cogzoo"
))
head(responses)
```

## Constructing and fitting a model

```{r}
mem <- data.frame(
  d1 = seq_len(nrow(responses)),
  d2 = rev(seq_len(nrow(responses))),
  category = rep(1:2, length.out = nrow(responses))
)
fit_data <- data.frame(
  d1 = mem$d1, d2 = mem$d2,
  n_cat1 = responses[["Cat-1-s"]],
  n_total = responses[["Cat-1-s"]] + responses[["Cat-2-s"]]
)

model <- gcm()
fitted <- cm_fit(
  model, fit_data, mem = mem,
  method = "L-BFGS-B", lower = c(0.01, 0.01, 0.01), upper = c(0.99, 5, 0.99)
)
fitted
```

## Predicting and checking calibration

```{r, fig.alt = "Scatterplot of observed vs. model-predicted category-1 response proportions, points falling roughly along the diagonal."}
predicted <- cm_predict(fitted, fit_data[c("d1", "d2")], mem = mem)
observed <- fit_data$n_cat1 / fit_data$n_total
plot(fitted, observed, predicted)
```

## Simulating

`cm_simulate()` draws synthetic data from a model — this is also what the
package's simulate-and-recover tests use to check a model's likelihood is
implemented correctly (see `CONTRIBUTING.md` for that pattern).

```{r}
cm_simulate(fitted, fit_data[c("d1", "d2")], mem = mem, n = 50)
```
